---
title: "Chapter 2 Notes"
author: "Tim"
date: "11/5/2017"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, 
                      results='show', cache=TRUE, autodep=TRUE)
```
## Estimation of the mean

\[ \mu_X = \frac{1}{N}\sum_{i=1}^n \hat{X_i}\]

## Estimate total population:

\[\hat{N} = \sum_{i=1}^n \frac{1}{\pi_i}\]

## Surveys in R

```{r}
library(tidyverse)
library(survey)
data(api)

ggplot(apipop, aes(x=enroll)) +
  geom_histogram()

samples <- 1:1000
schoolpop <- apipop$enroll[complete.cases(apipop$enroll)]

means <- apply(as.array(samples), 1, function(i) mean(sample(x=schoolpop, size = 200, replace = TRUE)))
d <- data.frame( sample = 1:1000)
d$means <- means
ggplot(data=d, aes(x = means)) +
  geom_histogram() + 
  geom_vline(xintercept = mean(schoolpop))
```

Next, we introduce some `survery` package tools. We create a survey design by looking at a random sample of California schools from the `api` data set. This is the a new perspective on the data above, estimated total population and mean from the random sample, based on a random sample of `r nrow(apisrs)`.

```{r}
srs_design <- svydesign(id=~1, fpc=~fpc, data=apisrs)
srs_design
```

```{r}
svytotal(~enroll, srs_design)
```

```{r}
svymean(~enroll, srs_design)
```

And here is the actual population and mean:

```{r}
sum(schoolpop)
mean(schoolpop)
```

Some extra details in the data. `pw` is the sampling weight and `fpc` is the finite population correction:


```{r}
unique(apisrs$pw)
unique(apisrs$fpc)
```

Where the sampling weight is 6194/200 = `r 6194/200`. As discussed in the text, with large samples, usually the case with complex surveys, the finite population correction `fpc` is very small. Here's an example:

```{r}
nofpc <- svydesign(id=~1, weights=~pw, data=apisrs)
svytotal(~enroll, nofpc)
svymean(~enroll, nofpc)
```

We only see a small increase in the standard error. 

On categorical type variables:

```{r}
svytotal(~stype, srs_design)
```

gives the estimated total population for elementary, middle, and high schools.

We can also call multiple variables to estimate, and use contrasts to make combinations:

```{r}
(means <- svymean(~api00+api99, srs_design))
svycontrast(means, c(api00=1, api99=-1))
svycontrast(means, quote(api00-api99))
```

Add variables with `update`:

```{r}
srs_design <- update(srs_design, apidiff=api00-api99)
svymean(~apidiff, srs_design)
```


## Stratified Sampling

The principle idea is to divide the population into groups called strata, with a separate probability sample from each one. Importantly, "a stratified sample is just a set of simple random samples from each stratum."

Let's take a look at a Stratified example that takes 100 elementary schools, 50 middle schools, and 50 high schools. Since high schools are the largest, a total random sample will include more high schools. The idea is to decrease the variance in groups with a smaller number. 

```{r}
strat_design <- svydesign(id=~1, strata = ~stype, fpc=~fpc, data=apistrat)
svytotal(~enroll, strat_design)
svymean(~enroll, strat_design)
```

The estimates have increased, meaning they are closer to the true values of `r mean(schoolpop)` and `r sum(schoolpop)`.

It is important to note a change in `fpc`:

```{r}
unique(apistrat$fpc)
```

The `fpc` corresponds to the total number of schools in each strata. Good quote from the book: "This extra information on the population is the source of the extra precision in the estimates, and the precision increase is greatest for the variables that are most accurately predicted by the population information."

For example, 
```{r}
svymean(~api00+api99, strat_design)
svytotal(~stype, strat_design)
```

`api` scores haven't improved much, since the breakdown by schools doesn't provide much more information. However now the population total is exact, since we specify the population of each subgroup.

## Replicate Weights

Replicate Weights are various methods used to calculate the standard error of a statistic, usually the ones besides total and mean.

Replicate weights compute the standard deviation across partially independent samples, and extrapolates this to the standard deviation between completely independent samples. 

```{r}
library(foreign)
chis_adult <- read.dta('Adult.dta')
```

If the survey specifies replicate weights directly, we can use those directly:

```{r}
chis <- svrepdesign(variables=chis_adult[,1:418],
                    repweights = chis_adult[,420:499],
                    weights=chis_adult[,419], combined.weights=TRUE,
                    type="other", scale=1, rscales=1)
```

If not given replicate weighs, we can create them:

```{r}
boot_design <- as.svrepdesign(strat_design, type="bootstrap", replicates=100)
jk_design <- as.svrepdesign(strat_design)

svymean(~enroll, boot_design)
svymean(~enroll, jk_design)
```

## Estimating subpopulations

Since each stratum is a random sample, to estimate statistics for that population we can only use the data from that state.

However, we are often concerned about other populations within the survey that do not make a stratum. For example, the 18-25 year old age group in BRFSS is not a stratum, meaning that the pairwise sampling probabilities are not necessarily non-zero, "because the number of participants from this age group is random rather than fixed in advance."

Here, "Since the sampling weights would be correct but the pairwise sampling probability would be incorrect, the resulting point estimates would be right, but the standard errors would be wrong."

Consider the `emer` variable in `api`:

```{r}
str(apipop$emer)
```

This is the proportion of teachers in the school with emergency teaching credentials. We can estimate the mean academic performance index of schools with varying needs for teachers:

```{r}
emerg_high <- subset(strat_design, emer>20)
emerg_low <- subset(strat_design, emer==0)

svymean(~api00+api99, emerg_high)
svymean(~api00+api99, emerg_low)

svytotal(~enroll, emerg_high)
svytotal(~enroll, emerg_low)
```

Unsurprisingly, schools with no emergency teachers tend to do much better on average based on `api` scores. Unfortunately there are many more schools that have greater than 20% emergency teachers. 

With the CHIS data, we can estimate BMI means for race and sex subgroups:

```{r}
svyby(~bmi_p, ~srsex+racehpr, svymean, design=chis, keep.names = FALSE)
```

Or by median, with a demonstration of ratios:

```{r}
medians <- svyby(~bmi_p, ~srsex+racehpr, svyquantile, design=chis, covmat=TRUE, 
      quantiles=0.5)
svycontrast(medians, quote(MALE.LATINO/FEMALE.LATINO))
```

## Exercises

#### 2.3 

We can compute all the cross-tabulations with by using `interaction`:
```{r}
ftable(svytotal(~interaction(ab29,ab30), chis)) %>%
  as.tibble() %>%
  spread(Var2, n) %>%
  filter(A!=0)
totals <- svyby(~interaction(ab29,ab30), ~racehpr+ins+marit, 
      svytotal, design=chis, keep.names=FALSE) 
totals %>%
  as.tibble() %>%
  select(racehpr, ins, marit, yes_yes = `interaction(ab29, ab30)YES.YES`)
```

Alternatively, we can directly specify the response which makes some things easier like rations. The output is also requires less tidying.

```{r}
svytotal(~I((ab29=='YES')*(ab30=='YES')), chis)
svytotal(~ab29+ab30, chis)
```

Using this syntax it is easy to specify ratios:

```{r hypertension_treat_ratio}
# ratio of treated to total cases
svyratio(~(ab30=="YES"), ~(ab29=="YES"), chis)
```
```{r}
svyby(~(ab30=="YES"), by=~ins,
      denominator=~(ab29=='YES'), design=chis, svyratio)
```

70% of people with insurance have treated hypertension, compared to only 30% of people without insurance.

```{r}
svyby(~(ab30=="YES"), by=~racehpr,
      denominator=~(ab29=='YES'), design=chis, svyratio)
```
```{r}
svyby(~(ab30=="YES"), by=~marit+ins,
      denominator=~(ab29=='YES'), design=chis, svyratio)
```
```{r}
svyby(~(ab30=="YES"), by=~racehpr+ins,
      denominator=~(ab29=='YES'), design=chis, svyratio)
```

Error: All replicates contained NAs. This means we don't have the data to fully estimate the subgroups broken down by race, insurance, and marriage. 

```{r}
# svyby(~(ab30=="YES"), by=~racehpr+ins+marit,
#       denominator=~(ab29=='YES'), design=chis, svyratio)
```

#### 2.4

We already did the first question in the previous exercise:

```{r, ref.label='hypertension_treat_ratio'}
```

Next, we want the proportion of diabetics `ab22` with diagnosed hypertension:

```{r}
svyratio(~(ab22=='YES'), ~(ab29=='YES'), chis)
```

And the proportion of hypertensive diabetics whose diabetes is being treated. The book say the variable for the treatment of diabetics, so I'll modify the question and consider the hypertensive diabetics whose hypertension is being treated:

```{r}
svyby(~(ab30=="YES"), by=~ab22,
      denominator=~(ab29=='YES'), design=chis, svyratio)
```

Given that some one has hypertension, they are more likely to receive treatment if they also have diabetes. 

#### 2.5





