---
title: "MRP Primer"
author: "Tim"
date: "11/9/2017"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, 
                      results='show', cache=TRUE, autodep=TRUE)
```
## Data and references

## Poststrat full example

```{r}
library(tidyverse)
library(lme4)
library(brms)
library(rstan)

rstan_options(auto_write=TRUE)
options(mc.cores=parallel::detectCores())
```
```{r}
marriage.data <- foreign::read.dta('gay_marriage_megapoll.dta', convert.underscore=TRUE)
```

## Data Cleaning

Kastellec does a lot of variable recoding and in some ways I'm not sure the tidy way is much better, but we'll try:

```{r, echo=FALSE}
Statelevel <- foreign::read.dta("state_level_update.dta",convert.underscore = TRUE)
Statelevel <- Statelevel[order(Statelevel$sstate.initnum),]

Census <- foreign::read.dta("poststratification 2000.dta",convert.underscore = TRUE)
Census <- Census[order(Census$cstate),]
Census$cstate.initnum <-  match(Census$cstate, Statelevel$sstate)

marriage.data$race.female <- (marriage.data$female *3) + marriage.data$race.wbh
marriage.data$age.edu.cat <- 4 * (marriage.data$age.cat -1) + marriage.data$edu.cat
marriage.data$p.evang.full <- Statelevel$p.evang[marriage.data$state.initnum]
marriage.data$p.mormon.full <-Statelevel$p.mormon[marriage.data$state.initnum]
marriage.data$p.relig.full <- marriage.data$p.evang.full + marriage.data$p.mormon.full
marriage.data$p.kerry.full <- Statelevel$kerry.04[marriage.data$state.initnum]

Census$crace.female <- (Census$cfemale *3) + Census$crace.WBH 
Census$cage.edu.cat <- 4 * (Census$cage.cat -1) + Census$cedu.cat 
Census$cp.evang.full<-  Statelevel$p.evang[Census$cstate.initnum]
Census$cp.mormon.full <- Statelevel$p.mormon[Census$cstate.initnum]
Census$cp.relig.full <- Census$cp.evang.full + Census$cp.mormon.full
Census$cp.kerry.full <-  Statelevel$kerry.04[Census$cstate.initnum]
```

```{r}
individual.model <- glmer(formula = yes.of.all ~ 
                            (1|race.female) + (1|age.cat) +
                            (1|edu.cat) + (1|age.edu.cat) + 
                            (1|state) + (1|region) + (1|poll) +
                            p.relig.full + p.kerry.full, 
                          data=marriage.data, family=binomial(link="logit"))
```

```{r}
#create vector of state ranefs and then fill in missing ones
state.ranefs <- array(NA,c(51,1))
dimnames(state.ranefs) <- list(c(Statelevel$sstate),"effect")
for(i in Statelevel$sstate){
    state.ranefs[i,1] <- ranef(individual.model)$state[i,1]
}
#set states with missing REs (b/c not in data) to zero
state.ranefs[,1][is.na(state.ranefs[,1])] <- 0
```

1. Each observation in the Census is a certain demographic group, identified by the following characteristics:

```{r, results='hold'}
Census %>% as.tibble() %>% 
  dplyr::select(crace.female, cage.cat, cedu.cat, cage.edu.cat, cstate)

Census %>% as.tibble() %>% 
  dplyr::select(cregion, cp.relig.full, cp.kerry.full)
```

2. Each observation's demographics are noted as a percent of the state's population

```{r}
Census %>% as.tibble() %>% 
  dplyr::select(cpercent.state)
```

3. For each observation's demographics, our model tells us their estimated support for gay marriage:

```{r}
cellpred <- invlogit(fixef(individual.model)["(Intercept)"] 
                     + ranef(individual.model)$race.female[Census$crace.female,1]
                     + ranef(individual.model)$age.cat[Census$cage.cat,1] 
                     + ranef(individual.model)$edu.cat[Census$cedu.cat,1]
                     + ranef(individual.model)$age.edu.cat[Census$cage.edu.cat,1] 
                     + state.ranefs[Census$cstate,1]   
                     + ranef(individual.model)$region[Census$cregion,1] 
                     + (fixef(individual.model)["p.relig.full"] *Census$cp.relig.full)
                     + (fixef(individual.model)["p.kerry.full"] *Census$cp.kerry.full)
)
str(cellpred)
```

4. Next, we take the weighted average of the proportion of support for gay marriage. 

```{r}                                    
cellpredweighted <- cellpred * Census$cpercent.state
```

5. Next, we average reach cell within each state to get the predicted state outcome: 

```{r}
statepred <- tibble(
  stateabv = unique(Census$cstate),
  pred.support = 100*as.vector(tapply(cellpredweighted,Census$cstate,sum))
)
```


