---
title: "Chapter 7 Notes"
author: "Tim"
date: "11/16/2017"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, 
                      results='show', cache=TRUE, autodep=TRUE)
```
## US Elections

```{r, results='hold'}
library(tidyverse)
library(survey)
data(api)

clus2_design <- svydesign(id=~dnum+snum, fpc=~fpc1+fpc2, data=apiclus2)

pop.types <- data.frame(stype=c("E","H","M"), Freq=c(4421,755,1018))
ps_design <- postStratify(clus2_design, strata=~stype, population=pop.types)

svytotal(~enroll, clus2_design, na.rm=TRUE)
svymean(~api00, clus2_design, na.rm=TRUE)
```
```{r, results='hold'}
svytotal(~enroll, ps_design, na.rm=TRUE)
svymean(~api00, ps_design, na.rm=TRUE)
```

And the actual results for the population:

```{r}
apipop %>% 
  summarise(total=sum(enroll, na.rm=TRUE), mean = mean(api00, na.rm=TRUE))
```

Note that the poststratified mean can be found just by applying the weights. Finding the standard error is the part that requires additional computation. Computation that a (multilevel) model could do. 

```{r}
# poststratification weight is total pop of strata divided by
# population sampled
ps_weight = c(4421/100, 755/50, 1018/50)
tot_pop = 6194

apiclus2 %>% as.tibble() %>%
  group_by(stype) %>%
  summarise(tot_enroll=sum(enroll, na.rm=TRUE), tot_api00=sum(api00, na.rm=TRUE)) %>%
  summarise(ps_tot_enroll=sum(tot_enroll*ps_weight), 
            ps_mean_api00=sum(tot_api00*ps_weight)/tot_pop)
```

It isn't perfect, but we also aren't trying to correct for clustering.

## Poststratification on Stratified Samples

Let's try the poststratification on stratified surveys:

```{r, results='hold'}
strat_design <- svydesign(id=~1, strata=~stype, fpc=~fpc, data=apistrat)
svytotal(~enroll, strat_design, na.rm=TRUE)
svymean(~api00, strat_design, na.rm=TRUE)
```

By hand: 

```{r}
apistrat %>% as.tibble() %>%
  group_by(stype) %>%
  summarise(tot_enroll=sum(enroll, na.rm=TRUE), tot_api00=sum(api00, na.rm=TRUE)) %>%
  summarise(ps_tot_enroll=sum(tot_enroll*ps_weight), 
            ps_mean_api00=sum(tot_api00*ps_weight)/tot_pop)
```

We get exactly the same answers.


